Supporting Information

### 2017-01-25 portfolio/figs/strategy-table.tex
Incorporate comments

% latex table generated in R 3.3.1 by xtable 1.8-2 package
Label & Description \\ 
  \midrule
Dun C & Dungeness crab \\ 
  Fin - ll & Finfish, longline, statewide \\ 
  Fin - otter & Finfish, otter trawl, statewide \\ 
  Fin - pot & Finfish, pot gear, statewide \\ 
  Hal - ll & Halibut, longline \\ 
  Herr - pound & Herring, pound \\ 
  Herr roe - gill & Herring roe, gillnet \\ 
  King C & King crab \\ 
  S - drift - CI, AP, BB (sock) & Salmon (mostly sockeye), drift gillnet, CI, AP, BB \\ 
  S - drift - PWS & Salmon, drift gillnet, PWS \\ 
  S - drift - SE & Salmon, drift gillnet, Southeast Alaska \\ 
  S - seine - AP & Salmon, purse seine, Alaska Penninsula \\ 
  S - seine - CH & Salmon, purse seine, Chignik \\ 
  S - seine - K & Salmon, purse seine, Cook Inlet and Kodiak \\ 
  S - seine - PWS & Salmon, purse seine, PWS \\ 
  S - seine - SE & Salmon, purse seine, Southeast Alaska \\ 
  S - set - NS/K & Salmon, set gillnet, Norton Sound and Kuskokwim \\ 
  S - set - PWS, CI, K, AP, BB (sock) & Salmon (mostly sockeye), set gillnet, PWS, CI, K,  AP, BB \\ 
  S - set - YAK & Salmon, set gillnet, Yakutat \\ 
  S - troll & Salmon, power troll, statewide \\ 
  Sab & Sablefish, longline, statewide \\ 
  Sea cuc & Sea cucumber \\ 
  Tan C & Tanner crab, pot gear \\ 
   \bottomrule

>>>>

% latex table generated in R 3.3.2 by xtable 1.8-2 package
Label & Description & Mean spp. diversity \\ 
  \midrule
Dun C & Dungeness crab & 1.0 \\ 
  King C & King crab & 1.0 \\ 
  Herr - pound & Herring, pound & 1.0 \\ 
  Sea cuc & Sea cucumber & 1.0 \\ 
  Tan C & Tanner crab, pot gear & 1.0 \\ 
  Hal - ll & Halibut, longline & 1.0 \\ 
  Fin - pot & Finfish, pot gear, statewide & 1.0 \\ 
  Herr roe - gill & Herring roe, gillnet & 1.0 \\ 
  Sab & Sablefish, longline, statewide & 1.0 \\ 
  S - drift - CI, AP, BB (sock) & Salmon (mostly sockeye), drift gillnet, CI, AP, BB & 1.1 \\ 
  Fin - ll & Finfish, longline, statewide & 1.1 \\ 
  S - set - PWS, CI, K, AP, BB (sock) & Salmon (mostly sockeye), set gillnet, PWS, CI, K,  AP, BB & 1.2 \\ 
  S - seine - CH & Salmon, purse seine, Chignik & 1.3 \\ 
  S - seine - PWS & Salmon, purse seine, PWS & 1.4 \\ 
  Fin - otter & Finfish, otter trawl, statewide & 1.5 \\ 
  S - set - YAK & Salmon, set gillnet, Yakutat & 1.7 \\ 
  S - troll & Salmon, power troll, statewide & 1.8 \\ 
  S - seine - K & Salmon, purse seine, Cook Inlet and Kodiak & 1.9 \\ 
  S - drift - PWS & Salmon, drift gillnet, PWS & 2.0 \\ 
  S - seine - SE & Salmon, purse seine, Southeast Alaska & 2.1 \\ 
  S - seine - AP & Salmon, purse seine, Alaska Penninsula & 2.2 \\ 
  S - set - NS/K & Salmon, set gillnet, Norton Sound and Kuskokwim & 2.5 \\ 
  S - drift - SE & Salmon, drift gillnet, Southeast Alaska & 2.5 \\ 
   \bottomrule

### 2017-01-25 portfolio/text/anderson-etal-rev-pfx.tex
Incorporate comments

\affil[d]{Alaska Department of Fish and Game,
Division of Commercial Fisheries,
PO BOX 115526,
Juneau, AK, USA 99811-5526, USA}

@@

% 250:
\begin{abstract}
Individuals relying on the extraction of natural resources for their
livelihood---fishers, farmers, miners, and loggers, among others---face high
levels of income variability driven by a mix of environmental, biological,
management, and
economic factors. Key to managing these industries is identifying
how regulatory actions and individual behaviour affect income variability and, by
extension, economic stability and the sustainable use of natural resources. In
commercial fisheries, communities and vessels fishing a greater diversity of
species have less revenue variability than those fishing fewer species.
However, it is unclear if these benefits extend to the actions of individual
fishers and how year-to-year changes in diversification affect revenue and
revenue variability. Here, we evaluate two axes by which fishers in Alaska,
can diversify fishing
activities. We show that, despite increasing specialization over the last 30
years, fishing a set of permits with higher species diversity reduces individual revenue
variability and fishing an additional permit is associated with higher revenue
and lower variability. However, increasing species diversity within the
constraints of existing permits has a fishery-dependent effect on revenue and is
usually (93\% probability) associated with increased revenue uncertainty
the following year. Our results demonstrate that the most effective option for
individuals to decrease revenue variability is to purchase permits to
participate in additional or more diverse fisheries. However, this option is
expensive, often limited by regulations such as license limitation or catch share programs, and
consequently unavailable to many individuals. With increasing environmental
variability in a changing climate, it will be particularly important that
individuals relying on natural resources for their livelihood have effective
strategies to reduce income variability.
 % (`r g2_dens`% probability)
\end{abstract}

@@

We use \(\log(R_{i,t-1}\)) as an offset, and thus our model describes the log
ratio of revenue from year-to-year, i.e., \(\log(R_{i,t}) - \log(R_{i,t-1}) =
  \log(R_{i,t}/R_{i,t-1})\). We model change in revenue because preliminary
analysis revealed non-stationary trends in revenue (resulting from experience
or other unmodelled factors). This approach eliminates having to deal with the
non-stationary time series and eliminates having to estimate thousands of
mean-revenue intercepts associated with individual fishers.
% TODO AH: Should you note that you tested for it?  You say that you assume stationarity after you differnece, but why don’t you test and report it?
We removed
person-year combinations without a subsequent year of revenue.
% TODO AH: what proportion of the data is this?

>>>>

\affil[d]{Division of Commercial Fisheries,
Alaska Department of Fish and Game, PO BOX 115526, Juneau, AK 99811, USA}

@@

% 250:
\begin{abstract}
Individuals relying on the extraction of natural resources for their
livelihood---fishers, farmers, miners, and loggers, among others---face high
levels of income variability driven by a mix of environmental, biological,
management, and
economic factors. Key to managing these industries is identifying
how regulatory actions and individual behaviour affect income variability and, by
extension, economic stability and the sustainable use of natural resources. In
commercial fisheries, communities and vessels fishing a greater diversity of
species have less revenue variability than those fishing fewer species.
However, it is unclear if these benefits extend to the actions of individual
fishers and how year-to-year changes in diversification affect revenue and
revenue variability. Here, we evaluate two axes by which fishers in Alaska,
can diversify fishing
activities. We show that, despite increasing specialization over the last 30
years, fishing a set of permits with higher species diversity reduces individual revenue
variability and fishing an additional permit is associated with higher revenue
and lower variability. However, increasing species diversity within the
constraints of existing permits has a fishery-dependent effect on revenue and is
usually (93\% probability) associated with increased revenue uncertainty
the following year. Our results demonstrate that the most effective option for
individuals to decrease revenue variability is to purchase permits to
participate in additional or more diverse fisheries. However, this option is
expensive, often limited by regulations such as license limitation or catch share programs, and
consequently unavailable to many individuals. With increasing climatic variability,
it will be particularly important that
individuals relying on natural resources for their livelihood have effective
strategies to reduce income variability.
 % (`r g2_dens`% probability)
\end{abstract}

@@

We use \(\log(R_{i,t-1}\)) as an offset, and thus our model describes the log
ratio of revenue from year-to-year, i.e., \(\log(R_{i,t}) - \log(R_{i,t-1}) =
  \log(R_{i,t}/R_{i,t-1})\). We model change in revenue because preliminary
analysis revealed non-stationary trends in revenue (resulting from experience
or other unmodelled factors). This approach eliminates having to deal with the
non-stationary time series and eliminates having to estimate thousands of
mean-revenue intercepts associated with individual fishers.
% TODO AH: Should you note that you tested for it?  You say that you assume stationarity after you differnece, but why don’t you test and report it? reduces...?
We removed
person-year combinations without a subsequent year of revenue.
% TODO AH: what proportion of the data is this?

### 2017-01-25 portfolio/text/ms.Rmd
Remaining edits
Incorporate comments

\dropcap{M}anaging natural resources and sustaining the livelihoods of those that depend
on their extraction is challenged by multiple sources of variability in the coupled human-natural system. These livelihoods tend to have high annual
variability in income relative to other professions [@mishra2002; @sethi2012b].
In addition to economic risk, such as changes in demand or prices, individuals
dependent on natural resources are also subject to biological and environmental
variability [@smith1986]. For example, drought and flooding are a major source
of risk for agricultural food security and farmers' incomes [@gregory2005], and
catastrophic disease outbreaks and wildfires increase risk for the logging
industry [@kirilenko2007].

@@

Individuals who rely on natural resources for income develop strategies to
reduce income variability. For example, farmers may rotate their crops to buffer
short-term climate variation and long-term environmental trends [@zentner1988].
Similarly, loggers maintain a diverse mix of tree species, age structure, and
harvest in multiple areas to buffer against disease outbreak, wildfires, and
price variation (REF). However, otherwise well-intentioned regulations may limit
how individuals diversify, or incentivize against diversification. For example,
in terrestrial agriculture, crop subsidies in the U.S. may incentivize some
farms to specialize (e.g. dairy, grains, soybeans; @mishra2002). In other
instances, market forces (e.g. prices or demand) may discourage the
adoption of diverse strategies and help promote specialization [@bradshaw2004].

Diversification may benefit commercial fishers, given the
extreme economic risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
obtainable by commercial fishers [@kasperski2013]. However, fishers maintain the
ability to lower risk by diversifying their fishing portfolio. They can, for
example, target as many species as allowed by a permit, own multiple
permits, or diversify spatially and fish in multiple regions. The effectiveness
of these diversification strategies for increasing revenue and reducing revenue
variability remains uncertain. For example, theory and empirical studies demonstrate that
there can be substantial benefits to specialization for both farmers and fishers
[@smith1986; @abdulai2000; @arnason2005]. Furthermore, the majority of research
examining the effects of diversification on revenue do so at aggregate levels
that combine individuals (e.g., counties or communities; [@kasperski2013;
@sethi2014; @cline2017]). Thus, key uncertainties remain about the relationship
among diversity, fishing revenue, and fishing revenue variability at the level
of the individual.

@@

<!-- TODO AH: I think our fisher apporach is good, but I wonder if it’s worth a footnote to address the potential critique that at the indivdual level we don’t observe their wages in non-fishing activities.  I’m not sure if this is worth bringing up, but this is always the case anyway with vessels where the skipper may do other things. -->

@@

\caption{Trends in revenue and diversification over time. (A) Map of fishing
regions in Alaska, USA. (B) Total revenue in billions of USD from `r
range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue) permit
holders. (C) Proportion of single permit holders for all fisheries (thick black
line) and with permit holders omitted each year who fished any of the `r
n_jacknifed` top species groups (thin lines). (D) Mean effective diversity of
species fished (Simpson's diversity index) for all fisheries (thick black line)
and with permit holders omitted each year who fished any of the `r n_jacknifed`
top species groups (thin lines). } \label{fig:ts} \end{center}
\end{figure}

@@

Comparing single permit holders to individuals holding that same permit plus
additional permits illustrates the expected reduction in revenue variability
from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
S\figspokeall). For example, individuals fishing a halibut permit plus
one additional permit always had lower expected revenue variability compared to
individuals fishing only a halibut permit (median reduction of `r
100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
\ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet strategy
substantially reduced expected revenue variability (Fig. \ref{fig:spokes}B).
Adding permits also usually increased expected revenue, although the permits
themselves represent a significant expense.

@@

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given strategy (set of permits) is more variable
(Fig. \ref{fig:within}). After accounting for fishing effort and changes in
annual average revenue for fishers with the same set of permits, specializing
tends to be associated with increased revenue but also increased variability
(Fig. \ref{fig:within}; Figs S\figfe, S\figre, S\ref{fig:noeffort}).
On average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
decrease in species diversity, and across strategies weighted equally, `r
g2_dens`% of the probability density supports a positive relationship between
specialization and revenue.

@@

Changing (increasing or decreasing) an individual's fished species diversity
within the constraints of a set of permits tends to be associated with greater
expected variability or uncertainty in year-to-year revenue (Fig.
\ref{fig:within} strategies above zero; Figs. S\figfe, S\figre,
S\figmirror). This effect was strongest for individuals with a halibut
permit or halibut and sablefish permits. Individuals with these strategies
experienced a ~2% increase in variability for a 1% change in the year-to-year
ratio of species diversity (Fig. \ref{fig:within}). Becoming more generalized in
species diversity also leads to increased revenue variability (Fig.
S\ref{fig:mirror}A), suggesting that the act of changing the number of species fished for a given set of permits is
a contributor to increased variability of revenue.

@@

Our findings across fishing strategies are in line with previous research at the
community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014; @cline2017], but our
findings within strategies, with respect to individual year-to-year behavior,
provide new insights. Here, specializing tended to be associated with increased
revenue. This could be for a number of reasons. First, specialization can
increase fishing efficiency [@smith1986]. Efficiency gains from specialization
can be seen in other natural resource sectors. For example, rice farmers in
Ghana can increase their efficiency by devoting more land and labor to rice
production compared to other crops [@abdulai2000]. Second, given the
availability of valuable resources, specializing can be highly profitable. For
instance, when cocoa prices have been high, Nigerian farmers have tended to
specialize on the crop [@mustapha1999] and specializing on valuable tobacco
farming can be the optimal strategy regardless of alternative crops
[@chavez2014]. In our analyses, fishers with the opportunity to fish highly
valuable sockeye salmon benefited the most from specializing. However,
specializing can have longterm risks if market, management, biological, or
environmental conditions change. For example, U.S. farmers specialized on wheat,
corn, or soybeans were hit harder than diversified farmers in the early 1980s
when a financial crisis hit [@ikerd1990].
<!-- TODO AB: We might want to include more examples / references from fisheries. -->
Third, diversifying --- a lack of
specialization --- can happen as a consequence of non-ideal circumstances such
as poor availability of a more valuable resource. For example, Ugandan
households relying on forestry resources for income increase their income
portfolio diversity following downward economic shocks [@debela2012].
Because specialization inherently holds risk from ``putting all one's eggs in one
basket,'' it is not suprising that a decline in the most valuable resource
would lead to a shift towards diversification.

@@

Our analysis provides a synthetic view of the effects of species diversity on
individuals over 30 years in one of the most productive fishing regions in the
world. However, this large temporal and geographical scope potentially obscures
time or regional-variation in the consequences of diversification on revenue and
revenue variability. Furthermore, in
addressing the overall pattern, we have averaged out the potential signal of several major discrete
events, such as the Exxon Valdez oil spill in 1989, that had significant
consequences for the ecological and social systems in Alaska [@paine1996]. In
addition to revenue or income variability, it is important to understand the
benefits of diversification in the context of rare and extreme disturbances
which occur in natural systems worldwide [@anderson2017].
<!-- TODO AH: Isn’t the variabiity you mention in this sentence part of this? -->

@@

\acknow{TODO. National Center for Ecological Analysis and Synthesis (NCEAS) working
group *Applying portfolio effects to the Gulf of Alaska ecosystem: Did
multi-scale diversity buffer against the Exxon Valdez oil spill?*. Other
participants... Smith...
}

>>>>

\dropcap{It} can be difficult for individuals to sustain a 
livelihood from natural resource extraction.
These livelihoods tend to have high annual
variability in income relative to other professions [@mishra2002; @sethi2012b].
In addition to economic risk, such as changes in demand or prices, individuals
dependent on natural resources are also subject to biological and environmental
variability [@smith1986]. For example, drought and flooding are a major source
of risk for agricultural food security and farmers' incomes [@gregory2005], and
catastrophic disease outbreaks and wildfires increase risk for the logging
industry [@kirilenko2007].
<!-- TODO Could add anthropogenic perturbations (e.g., Exxon Valdez) to this list or in the next paragraph. -->

@@

Individuals who rely on natural resources for income develop strategies to
reduce income variability. For example, farmers may TODO 
to buffer
short-term climate variation and long-term environmental trends [@zentner1988].
Similarly, loggers maintain a diverse mix of tree species, age structure, and
harvest in multiple areas to buffer against disease outbreak, wildfires, and
price variation (REF). However, otherwise well-intentioned regulations may limit
how individuals diversify, or incentivize against diversification. For example,
in terrestrial agriculture, crop subsidies in the U.S. may incentivize some
farms to specialize (e.g. dairy, grains, soybeans; @mishra2002). <!-- TODO LP: ? -->
In other
instances, market forces (e.g. prices or demand) may discourage the
adoption of diverse strategies and help promote specialization [@bradshaw2004].

Diversification may benefit commercial fishers, given the
extreme economic risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
obtainable by commercial fishers [@kasperski2013]. However, fishers maintain the
ability to lower risk by diversifying their fishing portfolio.
 <!-- TODO LP: this is the question you are trying to answer -->
They can, for
example, target as many species as allowed by a permit, own multiple
permits, or diversify spatially and fish in multiple regions. The effectiveness
of these diversification strategies for increasing revenue and reducing revenue
variability remains uncertain. For example, theory and empirical studies demonstrate that
there can be substantial benefits to specialization for both farmers and fishers
[@smith1986; @abdulai2000; @arnason2005]. Furthermore, the majority of research
examining the effects of diversification on revenue do so at aggregate levels
that combine individuals (e.g., counties or communities; [@kasperski2013;
@sethi2014; @cline2017]). Thus, key uncertainties remain about the relationship
among diversity, fishing revenue, and fishing revenue variability at the level
of the individual.

@@

<!-- TODO AH: I think our fisher apporach is good, but I wonder if it’s worth a footnote to address the potential critique that at the indivdual level we don’t observe their wages in non-fishing activities.  I’m not sure if this is worth bringing up, but this is always the case anyway with vessels where the skipper may do other things. triue, and for others-->

@@

\caption{Trends in revenue and diversification over time. (A) Map of major fishing
areas in Alaska, USA. (B) Total revenue in billions of USD from `r
range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue) permit
holders. (C) Proportion of single permit holders for all fisheries (thick black
line) and with permit holders omitted each year who fished any of the `r
n_jacknifed` top species groups (thin lines). (D) Mean effective diversity of
species fished (Simpson's diversity index) for all fisheries (thick black line)
and with permit holders omitted each year who fished any of the `r n_jacknifed`
top species groups (thin lines). } \label{fig:ts} \end{center}
\end{figure}

@@

Comparing single permit holders to individuals holding that same permit plus
additional permits illustrates the expected reduction in year-to-year revenue variability
from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
S\figspokeall). For example, individuals fishing a halibut permit plus
one additional permit always had lower expected revenue variability compared to
individuals fishing only a halibut permit (median reduction of `r
100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
\ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet strategy
substantially reduced expected revenue variability (Fig. \ref{fig:spokes}B).
Adding permits also usually increased expected revenue, although the permits
themselves represent a significant expense.

@@

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given strategy (set of permits) is more variable
(Fig. \ref{fig:within}). After accounting for fishing effort and changes in
annual average revenue for fishers with the same set of permits, specializing
tends to be associated with increased revenue but also increased variability
(Fig. \ref{fig:within}; Figs S\figfe, S\figre, S\ref{fig:noeffort}).
 <!-- TODO LP: effort is endogenous? -->
On average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
decrease in species diversity, and across strategies weighted equally, `r
g2_dens`% of the probability density supports a positive relationship between
specialization and revenue.

@@

Changing (increasing or decreasing) an individual's fished species diversity
within the constraints of a set of permits tends to be associated with greater
expected variability or uncertainty in year-to-year revenue (Fig.
\ref{fig:within} strategies above zero; Figs. S\figfe, S\figre,
S\figmirror). This effect was strongest for individuals with a halibut and sablefish permit. 
Individuals with this strategies
experienced a ~2% increase in variability for a 1% change in the year-to-year
ratio of species diversity (Fig. \ref{fig:within}).
Becoming more generalized in
species diversity also leads to increased revenue variability (Fig.
S\ref{fig:mirror}A), suggesting that the act of changing the number of species fished for a given set of permits is
a contributor to increased variability of revenue.

@@

Our findings across fishing strategies are in line with previous research at the
community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014; @cline2017], but our
findings within strategies, with respect to individual year-to-year behavior,
provide new insights. Here, specializing tended to be associated with increased
revenue. This could be for a number of reasons. First, specialization can
increase fishing efficiency [@smith1986]. Efficiency gains from specialization
can be seen in other natural resource sectors. For example, rice farmers in
Ghana can increase their efficiency by devoting more land and labor to rice
production compared to other crops [@abdulai2000]. <!-- TODO: weak -->
Second, given the
availability of valuable resources, specializing can be highly profitable. For
instance, when cocoa prices have been high, Nigerian farmers have tended to
specialize on the crop [@mustapha1999] and specializing on valuable tobacco
farming can be the optimal strategy regardless of alternative crops
[@chavez2014]. In our analyses, fishers with the opportunity to fish highly
valuable sockeye salmon benefited the most from specializing. However,
specializing can have longterm risks if market, management, biological, or
environmental conditions change. For example, U.S. farmers specialized on wheat,
corn, or soybeans were hit harder than diversified farmers in the early 1980s
when a financial crisis hit [@ikerd1990].
<!-- TODO AB: We might want to include more examples / references from fisheries. -->
Third, diversifying --- a lack of
specialization --- can happen as a consequence of non-ideal circumstances such
as poor availability of a more valuable resource. For example, Ugandan
households relying on forestry resources for income increase their income
portfolio diversity following downward economic shocks [@debela2012].
Because specialization inherently holds risk from ``putting all one's eggs in one
basket,'' it is not suprising that a decline in the most valuable resource
would lead to a shift towards diversification.

@@

Our analysis provides a synthetic view of the effects of species diversity on
individuals over 30 years in one of the most productive fishing regions in the
world. However, this large temporal and geographical scope potentially obscures
time or regional-variation in the consequences of diversification on revenue and
revenue variability. Furthermore, in
addressing the overall pattern, we have averaged out the potential 
signal of several major discrete
events, such as the *Exxon Valdez* oil spill in 1989, that was a significant
disruption to the management, ecological, 
and social structure of fisheries in Alaska [@paine1996]. In
addition to revenue or income variability, it is important to understand the
benefits of diversification in the context of rare and extreme disturbances
which occur in natural systems worldwide [@anderson2017].
<!-- TODO AH: Isn’t the variabiity you mention in this sentence part of this? -->

@@

\acknow{TODO. National Center for Ecological Analysis and Synthesis (NCEAS) working
group *Applying portfolio effects to the Gulf of Alaska ecosystem: Did
multi-scale diversity buffer against the Exxon Valdez oil spill?*. Other
participants... Smith... Support of Jordan Watson came from NOAA Fisheries' Spatial Economics Toolbox for Fisheries (FishSET) Project.
}

### 2017-01-25 portfolio/text/supp.Rnw
Incorporate comments

We then proceeded to combine similar permits in the following steps.
We combined a number of permits that were the same except for vessel size.
We combined permits for
otter trawl,
longline finfish,
pot gear.
king crab,
and
shrimp.
We grouped a number of permits that were targeting a single species
and differed only in region fished.
For example, someone fishing herring roe in Kodiak (G34K)
and someone fishing herring roe in Alaska Peninsula (G34M)
were both considered to have a herring roe permit strategy.
We combined permits across each of
dungeness crab,
herring roe purse seining,
herring roe gillnet,
herring hand picked,
herring pound fishing,
sea cucumber,
sea urchin, and
tanner crabs.

Combining salmon permits was more difficult because
the species composition varies widely based on geography.
We grouped salmon permits within gear type-region
combinations that shared similar species composition.
For the drift gillnet (S03) permits,
we combined permits from
Cook Inlet, Bristol Bay, and Alaska Peninsula because sockeye
represents greater than 90\% of species revenue.
For the purse seine permits (S01),
there was a wide variety in species revenue by
area ($\sim$ 85\% pink in Prince William Sound to
85\% sockeye on the Alaskan Peninsula).
Of the five permits, only two had similar species compositions,
so we grouped Kodiak and Cook Inlet.
For the set gillnet salmon permits (S04),
there was some variety but also clear sockeye specialists.
We grouped permits that had greater than 80\% sockeye revenue.
This combined permits across
Prince William Sound,
Cook Inlet,
Kodiak,
Alaska Peninsula,
and Bristol Bay.
We grouped permits from Norton Sound
and Kuskokwim because of similar species compositions.
The remaining salmon permits were relatively unique
and were left ungrouped.

@@

After this process of combining permits,
we were left with \Sexpr{n_grouped_permits}
unique fishing strategies (Table~\ref{tab:strategies}),
which we aggregated
within each person-year combination.
For example, if someone fished halibut
and sablefish in the same year
their strategy for that year would be `halibut-sablefish'.
We removed people who switched strategies from one year to the next
since we were working with changes in revenue from year to year
and so needed consistent strategies across any pair of years
for a permit holder.
We removed strategies with less than 100 unique permit holders
across all years to focus on the major strategies
leaving us with a total of \Sexpr{nstr} strategies.
Finally, we removed all person-year combinations without
revenue recorded in the following year
in order to calculate percent change in revenue
from year to year.
This reduced the rows (person-year combinations)
in the final dataset from \Sexpr{ndat$annual} to
\Sexpr{ndat$annual_diff}.

@@

\begin{table}[h!]
  \centering
  \caption{Labels and descriptions of fishing strategies. In many cases,
a particular strategy represents a group of similar permits across areas
or vessel sizes.}
  \label{tab:strategies}
  \begin{tabular}{ll}
  \toprule
\input{../figs/strategy-table.tex}
  \end{tabular}
\end{table}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1.06\textwidth]{../figs/stan-str-posteriors-dot.pdf}
\caption{Estimates of within-strategy effects from the hierarchical model.
Panels refer to parameters in Eqns \ref{eq:mean} and \ref{eq:cv}.
Shown are medians (dots) and 50\% and 95\% credible intervals
(thick and thin line segments).
Strategies are ordered from high to low mean
species diversity from top to bottom.}
\label{fig:re}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1\textwidth]{../figs/stan-gg-spoke-all.pdf}
\caption{The effect of adding a permit to one's strategy
on estimated revenue variability and median revenue.
This is an extended version of Fig.~2 that includes all major
strategy combinations.}
\label{fig:spoke-all}
\end{center}
\end{figure}

>>>>

We then proceeded to combine similar permits in the following steps.
We combined a number of permits that were the same except for vessel size.
We combined permits for
otter trawl,
longline finfish,
pot gear,
king crab,
and
shrimp.
We grouped a number of permits that were targeting a single species
and differed only in region fished.
For example, someone fishing herring roe in Kodiak (G34K)
and someone fishing herring roe in Alaska Peninsula (G34M)
were both considered to have a herring roe permit strategy.
We combined permits across each of
Dungeness crab,
herring roe purse seining,
herring roe gillnet,
herring hand picked,
herring pound fishing,
sea cucumber,
sea urchin, and
tanner crabs.

Combining salmon permits was more difficult because
the species composition varies widely based on geography.
We grouped salmon permits within gear type-region
combinations that shared similar species composition.
For the drift gillnet (S03) permits,
we combined permits from
Cook Inlet, Bristol Bay, and Alaska Peninsula because sockeye
represents greater than 90\% of species revenue.
For the purse seine permits (S01),
there was a wide variety in species revenue by
area ($\sim$ 85\% pink salmon in Prince William Sound to
85\% sockeye salmon on the Alaskan Peninsula).
Of the five permits, only two had similar species compositions,
so we grouped Kodiak and Cook Inlet.
For the set gillnet salmon permits (S04),
there was some variety but also clear sockeye specialists.
We grouped permits that had greater than 80\% sockeye revenue.
This combined permits across
Prince William Sound,
Cook Inlet,
Kodiak,
Alaska Peninsula,
and Bristol Bay.
We grouped permits from Norton Sound
and Kuskokwim because of similar species compositions.
The remaining salmon permits were relatively unique
and were left ungrouped.

@@

After this process of combining permits,
we were left with \Sexpr{n_grouped_permits}
unique fishing strategies (Table~\ref{tab:strategies}),
which we aggregated
within each person-year combination.
For example, if someone fished halibut
and sablefish in the same year
their strategy for that year would be `halibut-sablefish'.
We removed data for pairs of years in which people switched
strategies from one year to the next
since we were working with changes in revenue from year-to-year
and so needed consistent strategies across any pair of years
for a permit holder.
We removed strategies with less than 100 unique permit holders
across all years to focus on the major strategies
leaving us with a total of \Sexpr{nstr} (unique or combined) strategies.
Finally, we removed all person-year combinations without
revenue recorded in the following year
in order to calculate percent change in revenue
from year to year.
This reduced the rows (person-year combinations)
in the final dataset from \Sexpr{ndat$annual} to
\Sexpr{ndat$annual_diff}.

@@

\clearpage
% TODO: AB: I suggest 2 additional columns: 1) potential species that could be harvested under each permit strategy, 2) number of unique permits contained within each strategy. This would really help to show how much flexibility someone within each permit strategy might have.
\begin{table}[h!]
\footnotesize
  \centering
  \caption{Labels and descriptions of unique fishing strategies. In many cases,
a particular strategy represents a group of similar permits across areas
or vessel sizes.}
  \label{tab:strategies}
  \begin{tabular}{lll}
  \toprule
\input{../figs/strategy-table.tex}
  \end{tabular}
\end{table}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1.06\textwidth]{../figs/stan-str-posteriors-dot.pdf}
\caption{Estimates of within-strategy effects from the hierarchical model.
Panels refer to parameters in equations \ref{eq:mean} and \ref{eq:cv}.
Shown are medians (dots) and 50\% and 95\% credible intervals
(thick and thin line segments).
Strategies are ordered from high to low mean
species diversity from top to bottom.}
\label{fig:re}
\end{center}
\end{figure}

% TODO: tweak labels to fit
\begin{figure}[htbp]
\begin{center}
\includegraphics[width=1\textwidth]{../figs/stan-gg-spoke-all.pdf}
\caption{The effect of adding a permit to one's strategy
on estimated revenue variability and median revenue.
This is an extended version of Fig.~2 that includes all major
strategy combinations.}
\label{fig:spoke-all}
\end{center}
\end{figure}

% TODO: all figs referenced?
